### 2024-02-17 questionnaire.tex
Initial commit of very rough questionnaire

\documentclass[17pt]{extreport}
\usepackage{titlesec}
\titleformat*{\section}{\normalsize\bfseries}
\usepackage{charter}
\usepackage{latexsym}
\usepackage{geometry}
\geometry{a4paper, portrait, margin=0.5in}
\title{Requirements around space}
\date{}
\begin{document}
\thispagestyle{empty}

\section*{Do we want multiple rooms or would we settle for a single room?}
\begin{itemize}

\item\textbf{For}
  \begin{enumerate}
    \item Separate rooms can reduce noise
    \item Separate rooms can reduce mess e.g. workshop/electronics contaminating kitchen
  \end{enumerate}
\item\textbf{Against}
  \begin{enumerate}
    \item We may be unable to find a space of size at the start
    \item We may be able to find a space of size, albeit single open-plan room
    \item Multiple rooms may be expensive, and there may be noise requirements anyway which preclude a workshop area
  \end{enumerate}
\item\textbf{Preference}\\$\Box$ Must have multiple rooms $\Box$ Willing to settle for a single room
\end{itemize}

\section*{Maximum membership subscription}
The suggested sum is \pounds{25} as this is what other hackerspaces are charging.

$\Box$ Under \pounds{15} $\Box$ \pounds{15} $\Box$ \pounds{20} $\Box$ \pounds{25} $\Box$ Over \pounds{30}

\section*{Do we want a dedicated kitchen area?}
\begin{itemize}

\item\textbf{For}
  \begin{enumerate}
    \item Making food together may bring us together as a community
    \item Being able to prepare food makes the space more livable
    \item There are few spaces open late which are not pubs
  \end{enumerate}
\item\textbf{Against}
  \begin{enumerate}
    \item Communal food preparation implies additional communal cleaning
  \end{enumerate}
\item\textbf{Preference}\\$\Box$ Must have kitchen area $\Box$ Willing to settle for no kitchen area
\end{itemize}

\thispagestyle{empty}
\section*{Do we want a dedicated workshop area?}
\begin{itemize}

\item\textbf{For}
  \begin{enumerate}
    \item There have been tentative expressions of interest in woodworking/metalworking etc.
    \item Access to these tools is generally useful for projects
  \end{enumerate}
\item\textbf{Against}
  \begin{enumerate}
  \item Have to absolutely get on top of safety early on: significant risk of injury with heavy tools and lone working, i.e. large organisational/training burden
  \item Property may have significant noise requirements, e.g. after 10pm.
  \end{enumerate}
\item\textbf{Preference}\\$\Box$ Must have workshop area $\Box$ Willing to settle for no workshop area
\end{itemize}

\section*{Name of the space}
Do you have any suggestions for the name of the space, or any other comments?

\end{document}